---
title: "Small Area Estimation - UFH, MFH & Comparison Report"
author: "EU SAE Pipeline"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: cosmo
    self_contained: true
    fig_width: 10
    fig_height: 6
    md_extensions: "-smart"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(readxl)
  library(knitr)
  library(patchwork)
  library(sf)
  library(scales)
  library(here)
})

fig_dir   <- here::here("outputs", "figures")
data_dir  <- here::here("outputs", "data")
table_dir <- here::here("outputs", "tables")

`%||%` <- function(x, y) {
  if (is.null(x) || length(x) == 0) y else x
}

include_fig <- function(filename, ...) {
  path <- file.path(fig_dir, filename)
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    cat(sprintf("*Figure not available: %s*\n", filename))
  }
}

safe_read_csv <- function(filename, dir = table_dir) {
  path <- file.path(dir, filename)
  if (file.exists(path)) read.csv(path) else NULL
}
```

# Executive Summary

```{r load-comparison}
pov_fh  <- tryCatch(read_excel(file.path(data_dir, "pov_fh.xlsx")),
                     error = function(e) NULL)
pov_mfh <- tryCatch(read_excel(file.path(data_dir, "pov_mfh.xlsx")),
                     error = function(e) NULL)
pov_cmp <- tryCatch(read_excel(file.path(data_dir, "pov_comparison_detailed.xlsx")),
                     error = function(e) NULL)
mfh_artifacts <- tryCatch(readRDS(file.path(data_dir, "mfh_artifacts.rds")),
                          error = function(e) NULL)

sig_summary <- safe_read_csv("significance_summary.csv",
                              file.path(fig_dir, "tables"))
exec_summary <- safe_read_csv("executive_summary.csv",
                               file.path(fig_dir, "tables"))
```

```{r exec-summary-table}
if (!is.null(exec_summary)) {
  kable(exec_summary, digits = 4,
        caption = "Mean estimates and mean MSEs by year")
}
```

```{r sig-summary-table}
if (!is.null(sig_summary)) {
  kable(sig_summary, digits = 0,
        caption = "Number of domains with statistically significant changes")
}
```

# Comparison of UFH and MFH

This section compares the univariate (UFH) and multivariate (MFH) Fay-Herriot pipelines side by side, examining regression formulas, normality diagnostics, precision metrics, poverty maps, and change analyses.

```{r mfh-availability, results='asis'}
mfh_status <- if (!is.null(mfh_artifacts)) {
  list(
    fit_failed  = isTRUE(mfh_artifacts$model_fit_failed),
    refvar_zero = isTRUE(mfh_artifacts$refvar_zero),
    var_choice  = mfh_artifacts$var_choice %||% "(unset)",
    cov_choice  = mfh_artifacts$cov_choice %||% "(unset)",
    diag_model  = mfh_artifacts$diag_model %||% "(unset)",
    bad_rows    = mfh_artifacts$bad_vardir_rows %||% 0L,
    n_domains   = mfh_artifacts$n_domains %||% NA_integer_,
    n_fitted    = mfh_artifacts$n_domains_fitted %||% mfh_artifacts$n_domains %||% NA_integer_
  )
} else {
  NULL
}

mfh_not_executed <- !is.null(mfh_status) &&
  (mfh_status$fit_failed || mfh_status$refvar_zero)

mfh_change <- safe_read_csv("comparison_final.csv")
mfh_change_available <- !is.null(mfh_change) &&
  "diff" %in% names(mfh_change) &&
  any(is.finite(suppressWarnings(as.numeric(mfh_change$diff))))

if (!is.null(mfh_status) && mfh_status$bad_rows > 0 && !mfh_not_executed) {
  cat(sprintf(
    paste0(
      "<div class=\"alert alert-info\">",
      "<strong>Note:</strong> %s of %s domains had missing/non-finite sampling variances ",
      "and were excluded from MFH fitting. Their MFH estimates are reported as NA. ",
      "The MFH model was fitted on the remaining %s domains.",
      "</div>\n"
    ),
    mfh_status$bad_rows,
    mfh_status$n_domains,
    mfh_status$n_fitted
  ))
}

if (mfh_not_executed) {
  if (mfh_status$fit_failed) {
    reason <- "The selected MFH model did not return a valid fit object."
  } else {
    reason <- "The selected MFH model estimated the random-effects variance at zero, so no domain-specific shrinkage was produced."
  }

  cat(sprintf(
    paste0(
      "<div class=\"alert alert-warning\">",
      "<strong>MFH was not executed for this configuration.</strong><br>",
      "%s<br>",
      "Selected options: <code>var_choice = %s</code>, <code>cov_choice = %s</code>, ",
      "<code>diagnostic model = %s</code>.<br>",
      "MFH maps, UFH-vs-MFH comparisons, and MFH change analyses are omitted or NA-filled.",
      "</div>\n"
    ),
    reason,
    mfh_status$var_choice,
    mfh_status$cov_choice,
    mfh_status$diag_model
  ))
} else if (!is.null(mfh_status) && !mfh_change_available) {
  cat(
    paste0(
      "<div class=\"alert alert-info\">",
      "<strong>MFH fitted, but MFH change analysis is not available.</strong><br>",
      "The level estimates are available, but the MCPE-based change analysis did not produce finite differences. ",
      "This can happen with raw direct variances in the parametric bootstrap.",
      "</div>\n"
    )
  )
}
```

## Regression Models

```{r formula-table}
formulas <- safe_read_csv("model_formulas.csv",
                           file.path(fig_dir, "tables"))
if (!is.null(formulas)) {
  kable(formulas,
        caption = "Selected regression formulas used in the FH and MFH pipelines")
}
```

```{r coef-fh}
fh_coefs <- safe_read_csv("fh_coefficients.csv",
                           file.path(fig_dir, "tables"))
if (!is.null(fh_coefs)) {
  kable(fh_coefs, digits = c(0, 0, 0, 6, 6, 3, 4, 0),
        caption = "Estimated FH regression coefficients by year")
}
```

```{r coef-mfh}
mfh_coefs <- safe_read_csv("mfh_coefficients.csv",
                            file.path(fig_dir, "tables"))
if (!is.null(mfh_coefs)) {
  kable(mfh_coefs, digits = c(0, 0, 0, 6, 6, 3, 4, 0),
        caption = "Estimated MFH regression coefficients by year")
}
```

## Normality Diagnostics

```{r normality-tests}
normality <- safe_read_csv("normality_tests.csv",
                            file.path(fig_dir, "tables"))
if (!is.null(normality)) {
  kable(normality, digits = 4,
        caption = "Shapiro-Wilk normality tests for FH and MFH diagnostics")
}
```

### Q-Q Plots

```{r qq-plots, results='asis'}
qq_files <- list.files(file.path(fig_dir, "diagnostics"),
                       pattern = "^qq_.*\\.png$", full.names = FALSE)
if (length(qq_files) > 0) {
  for (f in sort(qq_files)) {
    label <- gsub("_", " ", tools::file_path_sans_ext(f))
    cat(sprintf("\n#### %s\n\n", label))
    cat(sprintf("![%s](%s)\n\n", label,
                file.path(fig_dir, "diagnostics", f)))
  }
} else {
  cat("*No Q-Q plots available.*\n")
}
```

### Density Plots

```{r density-plots, results='asis'}
density_files <- list.files(file.path(fig_dir, "diagnostics"),
                            pattern = "^density_.*\\.png$", full.names = FALSE)
if (length(density_files) > 0) {
  for (f in sort(density_files)) {
    label <- gsub("_", " ", tools::file_path_sans_ext(f))
    cat(sprintf("\n#### %s\n\n", label))
    cat(sprintf("![%s](%s)\n\n", label,
                file.path(fig_dir, "diagnostics", f)))
  }
} else {
  cat("*No density plots available.*\n")
}
```

## Rate, MSE, RMSE, and CV Comparisons

```{r metric-comparisons, results='asis'}
mc_files <- list.files(file.path(fig_dir, "metric_comparisons"),
                       pattern = "\\.png$", full.names = FALSE)
if (length(mc_files) > 0) {
  for (f in sort(mc_files)) {
    label <- gsub("_", " ", tools::file_path_sans_ext(f))
    cat(sprintf("\n### %s\n\n", label))
    cat(sprintf("![%s](%s)\n\n", label,
                file.path(fig_dir, "metric_comparisons", f)))
  }
} else {
  cat("*No metric comparison plots available.*\n")
}
```

## Poverty Maps

```{r poverty-maps, results='asis'}
map_files <- list.files(file.path(fig_dir, "poverty_maps"),
                        pattern = "\\.png$", full.names = FALSE)
if (length(map_files) > 0) {
  for (f in sort(map_files)) {
    label <- gsub("_", " ", tools::file_path_sans_ext(f))
    cat(sprintf("\n### %s\n\n", label))
    cat(sprintf("![%s](%s)\n\n", label,
                file.path(fig_dir, "poverty_maps", f)))
  }
} else {
  cat("*No poverty maps available.*\n")
}
```

## Statistical Significance of Poverty Changes

```{r sig-comparison}
sig_cmp <- tryCatch(
  read_excel(file.path(data_dir, "statistical_significance_comparison.xlsx")),
  error = function(e) NULL
)
if (!is.null(sig_cmp)) {
  kable(head(sig_cmp, 30), digits = 4,
        caption = "Statistical significance comparison (first 30 rows)")
}
```

```{r change-figures, results='asis'}
change_files <- list.files(file.path(fig_dir, "change_figures"),
                           pattern = "\\.png$", full.names = FALSE)
if (length(change_files) > 0) {
  for (f in sort(change_files)) {
    label <- gsub("_", " ", tools::file_path_sans_ext(f))
    cat(sprintf("\n### %s\n\n", label))
    cat(sprintf("![%s](%s)\n\n", label,
                file.path(fig_dir, "change_figures", f)))
  }
} else {
  cat("*No change figures available.*\n")
}
```

# Session Information

```{r session-info}
session_path <- here::here("outputs", "session_info.txt")
if (file.exists(session_path)) {
  cat("```\n")
  cat(readLines(session_path), sep = "\n")
  cat("\n```\n")
} else {
  cat("*Session info not yet generated.*\n")
}
```
